---
title: "EPI590R_Final"
author: "Monica Rivera"
format: html
editor: visual
echo: true
eval: true
---

## EPI 590R Final Project

Newborn Screening Disorders dataset: [**Screening Disorders by California Region**](https://catalog.data.gov/dataset/newborn-screening-disorders-426e4/resource/54bb2656-0073-4380-bbc7-7ac7919aee38 "Screening Disorders by California Region")

```{r}
# Load packages #
library(tidyverse)
library(gtsummary)

# Load data #
nsdca <- read_csv(here::here("data", "raw", "nsdca.csv"))
                  
# Clean column names #
nsdca <- nsdca %>%
  rename(
    region = `California Region`,
    disease_type = `Disease Type`,
    cases = `Case\nCount`,
    screened = `Number\nScreened`,
    percent_region = `Percent\nof All\nDisorders\nin Region`
  ) %>%

# Convert percent column to numeric & create factor for categorical variable: region #
  mutate(
    percent_region = as.numeric(percent_region),
    region = factor(region, labels = c("Los Angeles", "Central Coast", "Northern Central Valley", "Northern, Mountain, and Desert Areas", "San Francisco Bay Area", "South Coast", "Southern Inland Empire")
))

# Remove rows with missing values #
nsdca_cleaned <- nsdca %>%
  filter(!is.na(percent_region))

# Display the structure of the cleaned data frame #
str(nsdca_cleaned)

# Save the cleaned data to a new CSV file #
write_csv(nsdca_cleaned, "cleaned_nsdca.csv")

# Create data/clean folder #
if (!dir.exists(here::here("data", "clean"))) {
  dir.create(here::here("data", "clean"))
}

# Save the complete-case data #
write_rds(nsdca_cleaned, here::here("data", "clean", "nsdca_cleaned_complete.rds"))
```

1.  Create a `{gtsummary}` table of descriptive statistics about your data (1 pt)

```{r}
#| echo: false
2 * 2
```

2.  Fit a regression and present well-formatted results from the regression (1 pt)
3.  Create a figure (1 pt)
4.  Write *and* use a function that does something with the data (1 pt)
5.  The table, regression results, and figure, with appropriate captions (1 pt)
6.  Read in a dataset and save a file (can be data, table, figure, etc.) Use the `{here}` package every time you refer to file paths (at least twice) (1 pt)
7.  Commit and push your work to GitHub as you go (1 pt)
8.  In a README file, include any notes necessary for us to easily reproduce your analysis (e.g., “Run `script.R` and then render `document.qmd`”) as well as some information about your data (1 pt)
